2D Subquadratic Separable Distance Transformation for Path-Based Norms\thanks{This work has been mainly funded by ANR-11-BS02-009 and ANR-11-IDEX-0007-02 PALSE/2013/21 research grants.}

### 2014-05-14 latex/dgci2014/article.tex
Final

\begin{abstract}\sloppy
In many applications, separable algorithms have demonstrated their
efficiency to perform high performance and parallel volumetric
computations, such as distance transformation or medial axis
extraction. In the literature, several authors have discussed about
conditions on the metric to be considered in a separable approach.  In
this article, we present generic separable algorithms to efficiently
compute Voronoi maps and distance transformations for a large class of
metrics. Focusing to path based norms (chamfer masks, neighborhood
sequences, ...), we detail a subquadratic algorithm to compute such
volumetric transformation in dimension 2. More precisely, we describe
a $O(\log^2{m}\cdot N^2)$ algorithm for shapes in a $N\times N$ domain
with chamfer norm of size $m$.

@@

Since early works on digital geometry, distance transformation plays
an important role in many applications
\cite{Rosenfeld1966,Rosenfeld1968}. Given a finite input shape
$X\subset \Z^n$, the distance transformation labels each point in $X$
with the distance to its closest point in $\Z^n \setminus X$. Labeling
each point by the closest background point leads to Voronoi
maps. Since such characterization is parametrized by a distance
function, many authors address this distance transformation problem
with trade-offs between algorithmic performances and the
\emph{accuracy} of the digital distance function with respect to the
Euclidean one.  Hence, authors have considered distances based on
chamfer masks \cite{Rosenfeld1968,borgefors,fouard:ivc:2005} or
sequences of chamfer masks
\cite{Rosenfeld1966,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13};
the vector displacement based Euclidean distance
\cite{danielson,ragnemalm}; Voronoi diagram based Euclidean distance
\cite{BreuEtAl95,Maurer2003} or the square of the Euclidean distance
\cite{Hirata,roerdnik}.  For the Euclidean metric, separable
volumetric computations have demonstrated to be very efficient:
optimal $O(n\cdot N^n)$ time algorithms for shapes in $N^n$ domains,
optimal multi-thread/GPU implementation\ldots (please refer to
\cite{dcoeurjo_ChapDTWADGMM} for a discussion). For path-based metrics
(chamfer mask, -weighted- neighborhood sequences,\ldots), two main
techniques exist to compute the distance transformation. The first one
considers a weighted graph formulation of the problem and
Dijkstra-like algorithms on weighted graphs to compute distances. If
$m$ denotes the size of the chamfer mask, computational cost could be
in $O(m\cdot N^n)$ using a cyclic bucket data structure
\cite{verwer_uniform}. Another approach consists in a raster scan of
the domain: first the chamfer mask is decomposed into disjoint
sub-masks. Then the domain grid points are scanned in a given order
(consistent with the sub-mask construction) and a local computation is
performed before being propagated
\cite{Rosenfeld1966,borgefors}. Scanning the domain several times (one
per sub-mask) leads to the distance transformation values. Again, we
end up with a $O(m\cdot N^n)$ computational cost.

Beside specific applications which use the anisotropic nature of the
chamfer mask, rotational dependency is usually ensured with increasing
mask size $m$ together with weight optimization. In this context and
for arbitrarily large image $N$, both Dijsktra-like and raster scan
approach have computational cost which is quadratic with respect to
$N$ and $m$. In practical situations, $m \ll N$ but $m$ still needs to
be in $O(N^n)$ to have accurate asymptotic DT.

@@

metrics. Overall computational costs can be summarized as follows:

@@

(weighed, with $w_i\geq 0$) $L_p$ metrics

@@

defines metric spaces $(\Z^n, \R , d_{L_p})$ which are not

@@

  norms} which are chamfer metric induced by a norm).  To evaluate distances between two digital points for a

@@

  of $A$.}$ $f$ being the number of rows in $A$ and the number of

@@

distances can be obtained from similar expression to (\ref{eq:1}):

@@

computations are performed dimension by dimension

@@

In the following, we consider Voronoi map approach as defined in
\cite{BreuEtAl95}: Let us first define hyper-rectangular image $I_X:

@@

  same coordinates except at $i$.}. In this process, metric information

@@

$a,b,c\in\Z^n$ such that $a_i<b_i<c_i$ and a 1D image span
$S$\footnote{Subscript $a_i$ denotes the $i^{th}$ coordinate of point
  $a\in\Z^n$.}, this

@@

  \caption{Voronoi map construction on 1D image span $S$ along the
    $i^{th}$ dimension.\label{alg:voro}}
\end{algorithm}
\begin{figure}
  \begin{center}
    \subfigure[]{ \begin{overpic}[width=2cm]{figs/hiddenByTrue}
    \put(26,5){$a$}
    \put(0,35){$b$}
    \put(25,75){$c$}
    \put(55,-15){$S$}
    \end{overpic}}
    \subfigure[]{\begin{overpic}[width=2cm]{figs/hiddenByFalse}
    \put(37,8){$a$}
    \put(23,37){$b$}
    \put(30,72){$c$}
    \put(55,-15){$S$}
    \end{overpic}}~~~~~~~~~~
    \subfigure[]{\begin{overpic}[width=1.1cm]{figs/closest}
    \put(-5,52){$b$}
    \put(2,78){$c$}
    \put(47,55){$a$}
    \put(38,-15){$S$}
      \end{overpic}
    }
  \end{center}
  \caption{\small Geometrical predicates for Voronoi map construction:
    \textsc{HiddenBy}($a,b,c,S)$ returns false in $(a)$ and true in
    $(b)$ (straight segments correspond to Voronoi diagram
    edges). $(c)$ illustrates the \textsc{Closest}($a,b,c$) predicate
    for $c\in S$.}
  \label{fig:predicates}
\end{figure}

@@

For $L_p$ norms Algorithm \ref{alg:voro} provides exact Voronoi map
computation. Indeed, distance comparison predicate can be implemented
error-free from integer number comparisons considering the $p$ power of
the distance function $\left (d_{L_p}\right)^p$.

Proposition
\ref{prop:voronoi-map-from-1} also implies that most chamfer norms and
neighborhood sequence based norms can also be considered in separable
Algorithm \ref{alg:voro} (see Fig.~\ref{fig:metrics}). We just need
algorithmic tools to efficiently implement both \textsc{Closest} and
\textsc{HiddenBy} predicates.
\begin{figure}[htbp]
 {\includegraphics[width=1.65cm]{image-ball-1}}
 {\includegraphics[width=1.65cm]{image-ball-1-4}}
 {\includegraphics[width=1.65cm]{image-ball-2}}
 {\includegraphics[width=1.65cm]{image-ball-4}}
 {\includegraphics[width=1.65cm]{image-ball-43-1}}
 {\includegraphics[width=1.65cm]{image-ball-chamf34}}
 {\includegraphics[width=1.65cm]{image-ball-chamf5711}}
  \caption{Balls for different metrics satisfying Proposition \ref{prop:voronoi-map-from-1}: \emph{(from
      left to right)} $L_1$, $L_{1.4}$, $L_2$, $L_4$, $L_{43.1}$,
    $\M_{3-4}$ and $\M_{5-7-11}$.}
  \label{fig:metrics}
\end{figure}
%
\vspace{-1cm}
\subsection{A First Generic Adapter}
\label{sec:first-gener-adapt}
We first detail the overall computational cost of Algorithm
\ref{alg:voro}:
\begin{lemma}
\label{lem:generic}
Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
symmetric unit ball. If $C$ denotes the computational cost of
\textsc{Closest} predicate and $H$ is the computational cost of the
\textsc{HiddenBy} predicate, then Algorithm \ref{alg:voro} is in
$O(n\cdot N^n\cdot (C+H))$.
\end{lemma}
From \cite{BreuEtAl95,roerdnik}, $C=H=O(1)$ for $L_1,L_2$ and $L_\infty$
norms.  For prescribed norm $d$, we first define generic
Algorithms~\ref{alg:genclosest}, \ref{alg:genedge} and
\ref{alg:genhidden}: From some evaluations of $d$, \textsc{HiddenBy}
predicate is obtained by a dichotomic search on the 1D image span $S$
to localize the abscissa of Voronoi edges of sites $\{a,b\}$ and
$\{b,c\}$ (see Fig.~\ref{fig:algo}).

@@

\begin{algorithm}[H]\footnotesize
  $v_{ab}$ = \textsc{VoronoiEdge}$(a,b,s^1, s^{N_i})$\;
  $v_{bc}$ = \textsc{VoronoiEdge}$(b,c,s^1, s^{N_i})$\;
  \Return $(v_{ab}> v_{bc})$\;
  \caption{\footnotesize Generic \textsc{HiddenBy}($a,b,c\in\Z^n; S$ in the $i^{th}$
    direction) $a_i<b_i<c_i$.\label{alg:genhidden}}
\end{algorithm}
\begin{lemma}
\label{lem:genericDich}
   Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
   symmetric unit ball,   from Algorithms \ref{alg:genclosest} and
   \ref{alg:genhidden}, we have $H=O(C\cdot \log{N})$.
\end{lemma}
\begin{proof}
  First, the generic \textsc{VoronoiEdge} is dichotomic with
  $O(\log{N})$ steps and $O(C)$ for each one. \textsc{VoronoiEdge} is
  in $O(C\cdot\log{N})$.  To prove the correctness of
  \textsc{VoronoiEdge} (and thus Alg.~\ref{alg:voro}), we use the
  convexity of the metric and the quadrangle property: since
  $a_i<b_i$, all grid points closer to $a$ than $b$ in $S$ (if exist)
  are \emph{lower} than all pixels on $S$ closer to $b$ than to $a$
  (if exist too). Thanks to test in line $4$, recursive call maintain
  this invariant. Note that tests $2-5$ handle the fact that the edge
  may not belong to $S$.$\Box$
\end{proof}
If we consider a chamfer norm with $f$ facet rational ball, Eq.~(\ref{eq:1})
suggests that $C=O(f)$. Hence, we have the following corollary:
\begin{corollary}
  Let $\M$ be a chamfer norm whose rational ball has $f$ facets,
  separable exact Voronoi map $\Pi_X$ can be obtained in
    $O(n\cdot N^n\cdot f \cdot \log{N})$.
\end{corollary}
Please remember that naive implementation of chamfer mask distance
transformation using raster scan approach would lead to a $O(f\cdot
N^n)$ computational cost. In the following sections, we use the convex
structure of $\B$ to design a parallel subquadratic algorithm for chamfer
norms.

@@

Let us consider a 2D chamfer norm $\M$ with $m$ weighted vectors (note
that $f\EqDef|\B| = m$ in 2D). We suppose that vectors
$\{\vec{v}^k\}_{k=1\ldots m}$ are sorted counterclockwise. We define a wedge
as a pair $(\vec{v}^k,\vec{v}^{k+1})$ of vectors. To each wedge is
associated a row $A_k$ in the minimal H-representation of $A$ ($A_k$
can also be seen as a  --non-unitary-- normal vector to $\B$ facets
\cite{DBLP:journals/pr/NormandE09}). Using similar notations,
\cite{Thiel_hdr,Strand2008} demonstrate that the distance evaluation
of point $a$ can be obtained in two steps: First, we compute the wedge
$(\vec{v}^k,\vec{v}^{k+1})$, $a$ belongs to. Then,
\begin{equation}
\label{eq:2}
  d_\M(O,a) = A_k\cdot a^T
\end{equation}

@@

Please note that in practical implementations, we can use symmetries
in $\M$  to only work on restrictions of chamfer mask
directions, so called \emph{generator} in the literature.  To optimize the \textsc{HiddenBy} predicate, we focus on
the \textsc{VoronoiEdge} function. Given two points $a$ and $b$
($a_i<b_i$) and a 1D image span $S$ along the $i^{th}$ dimension, we
have to find the lowest abscissa $e_i$ of the point $e$ on $S$ such
that $d(a,e) < d(b,e)$ and $d(a,e')\geq d(b, e')$ for any $e'$ with
$e'_i>e_i$.
Let us first suppose that we do not have $e$ but we know the two wedges
$(\vec{v}^{k},\vec{v}^{k+1})$ (resp. $(\vec{v}^j,\vec{v}^{j+1})$) such
that the vector $(e-a)^T$ (resp. $(e-b)^T$) belongs to (see
Fig.~\ref{fig:algo}). In this situation, we know that $e$ is the
solution of
\begin{equation}
\label{eq:3}
  A_k\cdot  (e-a)^T = A_j\cdot (e -b)^T\;.
\end{equation}
(since $e\in S$, we have one linear equation with only one unknown
$e_i$). As a consequence, if we know the two wedges the Voronoi edge
belongs to, we have the abscissa in $O(1)$ (see Algorithm
\ref{alg:chamedge} and Fig.~\ref{fig:algo}$-(c)$).
\begin{algorithm}[h]\footnotesize
  $(\vec{v}^k,\vec{v}^{k+1})$ = \textsc{VoronoiEdgeWedge}$(a,b,\vec{v}^1,\vec{v}^m, S)$\;
  $(\vec{v}^j,\vec{v}^{j+1})$ = \textsc{VoronoiEdgeWedge}$(b,a,\vec{v}^1,\vec{v}^m, S)$\;

@@

  \Return $e_i$\;
  \caption{\footnotesize
2D chamfer norm \textsc{VoronoiEdge}($a,b,s^i,s^j\in\Z^2$).\label{alg:chamedge}}
\end{algorithm}
To obtain both wedges, we use a dichotomic search similar to
Algorithm~\ref{alg:genedge}: Algorithm~\ref{alg:chamedgew} returns the
wedge associated with $a$ containing the Voronoi edge with respect to
$b$. Applying this algorithm to obtain the wedge associated with $b$ with respect to
$a$ defines  Algorithm \ref{alg:chamedge}.
\begin{figure}[h]
  \begin{center}\scriptsize
    \subfigure[]{
      \begin{overpic}[width=2.5cm]{init}
        \put(-5,60){$a$}
        \put(85,40){$b$}
        \put(42,105){$S$}
      \end{overpic}
    }~~~~~~~~~
    \subfigure[]{
      \begin{overpic}[width=2.5cm]{algoMid}
      \put(-5,60){$a$}
        \put(85,40){$b$}
        \put(42,105){$S$}
        \put(12,60){$\vec{\vec{v}^{mid}}$}
        \put(13,80){$\vec{\vec{v}^{mid+1}}$}
        \put(-5,98){$\vec{\vec{v}^j}$}
        \put(5,32){$\vec{\vec{v}^i}$}
        \put(50,90){$p^{mid+1}$}
        \put(50,58){$p^{mid}$}
      \end{overpic}
    }~~~~~~~~~
    \subfigure[]{
      \begin{overpic}[width=2.5cm]{algoEnd}
      \put(-5,60){$a$}
        \put(85,40){$b$}
        \put(42,105){$S$}
        \put(50,50){$e$}
        \put(7,45){$\vec{\vec{v}^{k}}$}
        \put(11,70){$\vec{\vec{v}^{k+1}}$}
        \put(60,40){$\vec{\vec{v}^j}$}
        \put(67,57){$\vec{\vec{v}^{j+1}}$}
      \end{overpic}
    }
  \end{center}
  \caption{\small \textsc{VoronoiEdgeWedge} and \textsc{VoronoiEdge}: $(a)$
    initial problem, we want to compute the intersection between $S$
    and the Voronoi edge of $a$ and $b$ (in red). $(b)$ an internal step of
    \textsc{VoronoiEdgeWedge} to reduce the set of directions of $\M$
    at $a$ (here the next recursive call will be on
    $(\vec{v}^i,\vec{v}^{mid})$).$(c)$ final step of
    \textsc{VoronoiEdge} where both wedges have been obtained and thus
    $e$ can be computed.}
  \label{fig:algo}
\end{figure}
The dichotomic search shrinks the set of vectors
$\{\vec{v}^i,\ldots,\vec{v}^j\}$ to end up with a wedge
$(\vec{v}^k,\vec{v}^{k+1})$ such that the intersection point between
the straight line $(a + \vec{v}^k)$ and $S$ is in the Voronoi cell of
$a$ and such that the intersection between $(a + \vec{v}^k)$ and $S$
is in the Voronoi cell of $b$ (see
Fig.~\ref{fig:algo}$-(c)$). Algorithm~\ref{alg:chamedgew} thus first
compute the intersection points associated with a wedge
$(\vec{v}^{i+(j-i)/2}, \vec{v}^{i+(j-i)/2+1})$ (lines $5-6$);
Evaluates the distances at these points (lines $7-8$) and then decides
which set $\{\vec{v}^i,\ldots,\vec{v}^{i+(j-i)/2}\}$ or
$\{\vec{v}^{i+(j-i)/2},\ldots,\vec{v}^j\}$ has to be considered for
the recursive call (lines $14-25$ and Fig.~\ref{fig:algo}$-(b)$).

@@

In this section, we give some implementation details and experimental
results for chamfer norm Voronoi map in dimension 2.  First of all,
most algorithms presented here are available in the \textsc{DGtal}
library \cite{dgtal}. For $L_p$ metrics, we have implemented several
\textsc{Closest} and \textsc{HiddenBy} predicates: If $p=\{1,2\}$,
exact computations are proposed and all predicates are in $O(1)$ with
only integer number computations
\cite{Hirata,Maurer2003,roerdnik}; If $p\in \R$, $p\geq 1$, we
have approximated computations on real numbers (\emph{double}) and we
consider the Generic \textsc{HiddenBy} predicate in $O(\log N)$
(Alg.~\ref{alg:genhidden}). Since predicates are based on floating
point computations, numerical issues may occur. If $p\in \Z$, $p\geq
3$, we use exact integer number based computations of distances
storing sum of power $p$ quantities (which can be computed in
$O(\log{p})$ thanks to exponentiation by squaring). The
\textsc{HiddenBy} predicate is also based on
Algorithm~\ref{alg:genhidden}.
Beside these predicates for $L_p$ metrics, \textsc{DGtal} also
contains a generic metric adapter: if the user specifies a distance
function (taking two points and returning a value) corresponding to a
norm with axis symmetric unit ball, generic \textsc{Closest} and \textsc{HiddenBy} predicates can be
automatically constructed. Please note that since all algorithms are
separable, generic framework provided in \textsc{DGtal} allow us to
have a free multi-thread implementation \cite{dcoeurjo_ChapDTWADGMM}.

>>>>

\begin{abstract}\sloppy
In many applications, separable algorithms have demonstrated their
efficiency to perform high performance and parallel volumetric
computations, such as distance transformation or medial axis
extraction. In the literature, several authors have discussed about
conditions on the metric to be considered in a separable approach.  In
this article, we present generic separable algorithms to efficiently
compute Voronoi maps and distance transformations for a large class of
metrics. Focusing on path based norms (chamfer masks, neighborhood
sequences, ...), we detail a subquadratic algorithm to compute such
volumetric transformation in dimension 2. More precisely, we describe
a $O(\log^2{m}\cdot N^2)$ algorithm for shapes in a $N\times N$ domain
with chamfer norm of size $m$.

@@

Since early works on digital geometry, distance transformation has
been playing
an important role in many applications
\cite{Rosenfeld1966,Rosenfeld1968}. Given a finite input shape
$X\subset \Z^n$, the distance transformation labels each point in $X$
with the distance to its closest point in $\Z^n \setminus X$. Labeling
each point by the closest background point leads to Voronoi
maps. Since such characterization is parametrized by a distance
function, many authors address this distance transformation problem
with trade-offs between algorithmic performances and the
\emph{accuracy} of the digital distance function with respect to the
Euclidean one.  Hence, authors have considered distances based on
chamfer masks \cite{Rosenfeld1968,borgefors,fouard:ivc:2005} or
sequences of chamfer masks
\cite{Rosenfeld1966,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13};
the vector displacement based Euclidean distance
\cite{danielson,ragnemalm}; Voronoi diagram based Euclidean distance
\cite{BreuEtAl95,Maurer2003} or the square of the Euclidean distance
\cite{Hirata,roerdnik}.  For the Euclidean metric, separable
volumetric computations have demonstrated to be very efficient:
optimal $O(n\cdot N^n)$ time algorithms for shapes in $N^n$ domains,
optimal multi-thread/GPU implementation\ldots (please refer to
\cite{dcoeurjo_ChapDTWADGMM} for a discussion). For path-based metrics
(chamfer mask, -weighted- neighborhood sequences,\ldots), two main
techniques exist to compute the distance transformation. The first one
considers a weighted graph formulation of the problem and
Dijkstra-like algorithms on weighted graphs to compute distances. If
$m$ denotes the size of the chamfer mask, computational cost could be
in $O(m\cdot N^n)$ using a cyclic bucket data structure
\cite{verwer_uniform}. Another approach consists in a raster scan of
the domain: first the chamfer mask is decomposed into disjoint
sub-masks. Then the domain grid points are scanned in a given order
(consistent with the sub-mask construction) and a local computation is
performed before being propagated
\cite{Rosenfeld1966,borgefors}. Scanning the domain several times (one
per sub-mask) leads to the distance transformation values. Again, we
end up with a $O(m\cdot N^n)$ computational cost.

Beside specific applications which use the anisotropic nature of the
chamfer mask, rotational dependency is usually ensured  increasing
the mask size $m$ together with optimizing weights. In this context and
for arbitrarily large  $N$, both Dijsktra-like and raster scan
approaches have a quadratic computational cost   with respect to
$N$ and $m$. In practical situations, $m \ll N$ but $m$ still needs to
be in $O(N^n)$ to have accurate asymptotic DT.

@@

metrics. Overall computational costs can be summarized as follows (see
\ref{sec:first-gener-adapt} for predicate definitions):

@@

(weighted, with $w_i\geq 0$) $L_p$ metrics

@@

define metric spaces $(\Z^n, \R , d_{L_p})$ which are not

@@

  norms} which are chamfer metric induced by a norm.  To evaluate distances between two digital points for a

@@

  of $A$.}$ $f$ is the number of rows in $A$ and the number of

@@

distances can be obtained from a expression similar to (\ref{eq:1}):

@@

computations are performed dimension by dimension.

@@

In the following, we consider \emph{Voronoi Map} approach as defined in
\cite{BreuEtAl95}: Let us first define an hyper-rectangular image $I_X:

@@

  same coordinates except at their $i^{th}$ one.}. In this process, metric information

@@

$a,b,c\in\Z^n$ such that $a_i<b_i<c_i$\footnote{Subscript $a_i$ denotes the $i^{th}$ coordinate of point
  $a\in\Z^n$.} and a 1D image span
$S$, this

@@

  \caption{Voronoi map construction on 1D image span $S$ along the
    $i^{th}$ dimension.\label{alg:voro}}
\end{algorithm}
\begin{figure}
  \begin{center}
    \subfigure[]{ \begin{overpic}[width=2cm]{figs/hiddenByTrue}
    \put(26,5){$a$}
    \put(0,35){$b$}
    \put(25,75){$c$}
    \put(55,-15){$S$}
    \end{overpic}}
    \subfigure[]{\begin{overpic}[width=2cm]{figs/hiddenByFalse}
    \put(37,8){$a$}
    \put(23,37){$b$}
    \put(30,72){$c$}
    \put(55,-15){$S$}
    \end{overpic}}~~~~~~~~~~
    \subfigure[]{\begin{overpic}[width=1.1cm]{figs/closest}
    \put(-5,52){$b$}
    \put(2,78){$c$}
    \put(47,55){$a$}
    \put(38,-15){$S$}
      \end{overpic}
    }
  \end{center}
  \caption{\small Geometrical predicates for Voronoi map construction:
    \textsc{HiddenBy}($a,b,c,S)$ returns true in $(a)$ and false in
    $(b)$ (straight segments correspond to Voronoi diagram
    edges). $(c)$ illustrates the \textsc{Closest}($a,b,c$) predicate
    for $c\in S$.}
  \label{fig:predicates}
\end{figure}

@@

For $L_p$ norms Algorithm \ref{alg:voro} provides exact Voronoi map
computation. Indeed, distance comparison predicate can be error-free implemented
 from integer number comparisons considering the $p$ power of
the distance function $\left (d_{L_p}\right)^p$.

Proposition
\ref{prop:voronoi-map-from-1} also implies that most chamfer norms and
neighborhood sequence based norms can also be considered in separable
Algorithm \ref{alg:voro} (see Fig.~\ref{fig:metrics}). We just need
algorithmic tools to efficiently implement both \textsc{Closest} and
\textsc{HiddenBy} predicates.
\begin{figure}[htbp]
 {\includegraphics[width=1.65cm]{image-ball-1}}
 {\includegraphics[width=1.65cm]{image-ball-1-4}}
 {\includegraphics[width=1.65cm]{image-ball-2}}
 {\includegraphics[width=1.65cm]{image-ball-4}}
 {\includegraphics[width=1.65cm]{image-ball-43-1}}
 {\includegraphics[width=1.65cm]{image-ball-chamf34}}
 {\includegraphics[width=1.65cm]{image-ball-chamf5711}}
  \caption{Balls for different metrics satisfying Proposition \ref{prop:voronoi-map-from-1}: \emph{(from
      left to right)} $L_1$, $L_{1.4}$, $L_2$, $L_4$, $L_{43.1}$,
    $\M_{3-4}$ and $\M_{5-7-11}$.}
  \label{fig:metrics}
\end{figure}
%
\vspace{-1cm}
\subsection{A First Generic Adapter}
\label{sec:first-gener-adapt}
We first detail the overall computational cost of Algorithm
\ref{alg:voro}:
\begin{lemma}
\label{lem:generic}
Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
symmetric unit ball. If $C$ denotes the computational cost of
\textsc{Closest} predicate and $H$ is the computational cost of the
\textsc{HiddenBy} predicate, then Algorithm \ref{alg:voro} is in
$O(n\cdot N^n\cdot (C+H))$.
\end{lemma}
From \cite{BreuEtAl95,roerdnik}, $C=H=O(1)$ for $L_1,L_2$ and $L_\infty$
norms.  For given norm $d$, we first define generic
Algorithms~\ref{alg:genclosest}, \ref{alg:genedge} and
\ref{alg:genhidden}: From some evaluations of $d$, \textsc{HiddenBy}
predicate is obtained by a dichotomic search on the 1D image span $S$
to localize the abscissa of Voronoi edges of sites $\{a,b\}$ and
$\{b,c\}$ (see Fig.~\ref{fig:algo}).

@@

\begin{algorithm}[H]\footnotesize
  $v_{ab}$ = \textsc{VoronoiEdge}$(a,b,s^1, s^{N_i})$\;
  $v_{bc}$ = \textsc{VoronoiEdge}$(b,c,s^1, s^{N_i})$\;
  \Return $(v_{ab}> v_{bc})$\;
  \caption{\footnotesize Generic \textsc{HiddenBy}($a,b,c\in\Z^n; S$ in the $i^{th}$
    direction) $a_i<b_i<c_i$.\label{alg:genhidden}}
\end{algorithm}
\begin{lemma}
\label{lem:genericDich}
   Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
   symmetric unit ball,   from Algorithms \ref{alg:genclosest} and
   \ref{alg:genhidden}, we have $H=O(C\cdot \log{N})$.
\end{lemma}
\begin{proof}
  First, the generic \textsc{VoronoiEdge} is dichotomic with
  $O(\log{N})$ steps and each step is in  $O(C)$. \textsc{VoronoiEdge} is
  in $O(C\cdot\log{N})$.  To prove the correctness of
  \textsc{VoronoiEdge} (and thus Alg.~\ref{alg:voro}), we use the
  convexity of the metric and the quadrangle property: since
  $a_i<b_i$, all grid points closer to $a$ than $b$ in $S$ (if exist)
  are \emph{lower} than all pixels on $S$ closer to $b$ than to $a$
  (if exist too). Thanks to the test in line $8$, recursive call maintain
  this invariant. Note that tests on lines $2-5$ handle the fact that the edge
  may not belong to $S$.$\Box$
\end{proof}
If we consider a chamfer norm with a rational ball of $f$ facets, Eq.~(\ref{eq:1})
suggests that $C=O(f)$. Hence, we have the following corollary:
\begin{corollary}
  Let $\M$ be a chamfer norm whose rational ball has $f$ facets,
  separable exact Voronoi map $\Pi_X$ can be obtained in
    $O(n\cdot N^n\cdot f \cdot \log{N})$.
\end{corollary}
Please remember that naive implementation of chamfer mask distance
transformation using raster scan approach would lead to a $O(f\cdot
N^n)$ computational cost. In the following sections, we use the convex
structure of $\B$ to design a parallel subquadratic algorithm for chamfer
norms.

@@

Let us consider a 2D chamfer norm $\M$ with $m$ weighted vectors (note
that $f\EqDef|\B| = m$ in 2D). We suppose that vectors
$\{\vec{v}^k\}_{k=1\ldots m}$ are sorted counterclockwise. We define a wedge
as a pair $(\vec{v}^k,\vec{v}^{k+1})$ of vectors. To each wedge is
associated a row $A_k$ in the minimal H-representation of $A$ ($A_k$
can also be seen as a  --non-unitary-- normal vector to $\B$ facets
\cite{DBLP:journals/pr/NormandE09}). Using similar notations,
\cite{Thiel_hdr,Strand2008} demonstrate that the distance evaluation
of point $a$ can be obtained in two steps: First, we compute the wedge
$(\vec{v}^k,\vec{v}^{k+1})$ $a$ belongs to. Then,
\begin{equation}
\label{eq:2}
  d_\M(O,a) = A_k\cdot a^T
\end{equation}

@@

Please note that in practical implementations, we can use symmetries
in $\M$  to only work on restrictions of chamfer mask
directions, so called \emph{generator} in the literature.  To optimize the \textsc{HiddenBy} predicate, we focus on
the \textsc{VoronoiEdge} function. Given two points $a$ and $b$
($a_i<b_i$) and a 1D image span $S$ along the $i^{th}$ dimension, we
have to find the lowest abscissa $e_i$ of the point $e$ on $S$ such
that $d(a,e) < d(b,e)$ and $d(a,e')\geq d(b, e')$ for any $e'$ with
$e'_i>e_i$.
Let us first suppose that we do not know $e$ but we know the  wedge
$(\vec{v}^{k},\vec{v}^{k+1})$ (resp. $(\vec{v}^j,\vec{v}^{j+1})$) to which the vector $(e-a)^T$ (resp. $(e-b)^T$) belongs to (see
Fig.~\ref{fig:algo}$-(c)$). In this situation, we know that $e$ is the
solution of
\begin{equation}
\label{eq:3}
  A_k\cdot  (e-a)^T = A_j\cdot (e -b)^T\;.
\end{equation}
(since $e\in S$, we have one linear equation with only one unknown
$e_i$). As a consequence, if we know the two wedges the Voronoi edge
belongs to, we have the abscissa in $O(1)$ (see Algorithm
\ref{alg:chamedge} and Fig.~\ref{fig:algo}$-(c)$).
\begin{algorithm}[h]\footnotesize
  $(\vec{v}^k,\vec{v}^{k+1})$ = \textsc{VoronoiEdgeWedge}$(a,b,\vec{v}^1,\vec{v}^m, S)$\;
  $(\vec{v}^j,\vec{v}^{j+1})$ = \textsc{VoronoiEdgeWedge}$(b,a,\vec{v}^1,\vec{v}^m, S)$\;

@@

  \Return $e_i$\;
  \caption{\footnotesize
2D chamfer norm \textsc{VoronoiEdge}($a,b,s^i,s^j\in\Z^2$).\label{alg:chamedge}}
\end{algorithm}
To obtain both wedges, we use a dichotomic search similar to
Algorithm~\ref{alg:genedge}: Algorithm~\ref{alg:chamedgew} returns the
wedge associated with $a$ containing the Voronoi edge with respect to
$b$. Applying this algorithm to obtain the wedge associated with $b$ with respect to
$a$ defines  Algorithm \ref{alg:chamedge}.
\begin{figure}[h]
  \begin{center}\scriptsize
    \subfigure[]{
      \begin{overpic}[width=2.5cm]{init}
        \put(-5,60){$a$}
        \put(85,40){$b$}
        \put(42,105){$S$}
      \end{overpic}
    }~~~~~~~~~
    \subfigure[]{
      \begin{overpic}[width=2.5cm]{algoMid}
      \put(-5,60){$a$}
        \put(85,40){$b$}
        \put(42,105){$S$}
        \put(12,60){$\vec{\vec{v}^{mid}}$}
        \put(13,80){$\vec{\vec{v}^{mid+1}}$}
        \put(-5,98){$\vec{\vec{v}^j}$}
        \put(5,32){$\vec{\vec{v}^i}$}
        \put(50,90){$p^{mid+1}$}
        \put(50,58){$p^{mid}$}
      \end{overpic}
    }~~~~~~~~~
    \subfigure[]{
      \begin{overpic}[width=2.5cm]{algoEnd}
      \put(-5,60){$a$}
        \put(85,40){$b$}
        \put(42,105){$S$}
        \put(50,50){$e$}
        \put(7,45){$\vec{\vec{v}^{k}}$}
        \put(11,70){$\vec{\vec{v}^{k+1}}$}
        \put(60,40){$\vec{\vec{v}^j}$}
        \put(67,57){$\vec{\vec{v}^{j+1}}$}
      \end{overpic}
    }
  \end{center}
  \caption{\small \textsc{VoronoiEdgeWedge} and \textsc{VoronoiEdge}: $(a)$
    initial problem, we want to compute the intersection between $S$
    and the Voronoi edge of $a$ and $b$ (in red). $(b)$ an internal step of
    \textsc{VoronoiEdgeWedge} to reduce the set of directions of $\M$
    at $a$ (here the next recursive call will be on
    $(\vec{v}^i,\vec{v}^{mid})$).$(c)$ final step of
    \textsc{VoronoiEdge} where both wedges have been obtained and thus
    $e$ can be computed.}
  \label{fig:algo}
\end{figure}
The dichotomic search shrinks the set of vectors
$\{\vec{v}^i,\ldots,\vec{v}^j\}$ to end up with a wedge
$(\vec{v}^k,\vec{v}^{k+1})$ such that the intersection point between
the straight line $(a + \vec{v}^k)$ and $S$ is in the Voronoi cell of
$a$ and such that the intersection between $(a + \vec{v}^{k+1})$ and $S$
is in the Voronoi cell of $b$ (see
Fig.~\ref{fig:algo}$-(c)$). Algorithm~\ref{alg:chamedgew} thus first
computes the intersection points associated with a wedge
$(\vec{v}^{i+(j-i)/2}, \vec{v}^{i+(j-i)/2+1})$ (lines $5-6$);
evaluates the distances at these points (lines $7-8$) and then decides
which set $\{\vec{v}^i,\ldots,\vec{v}^{i+(j-i)/2}\}$ or
$\{\vec{v}^{i+(j-i)/2},\ldots,\vec{v}^j\}$ has to be considered for
the recursive call (lines $14-20$ and Fig.~\ref{fig:algo}$-(b)$).

@@

In this section, we give some implementation details and experimental
results for chamfer norm Voronoi map in dimension 2.  First of all,
most algorithms presented here are available in the \textsc{DGtal}
library \cite{dgtal}. For $L_p$ metrics, we have implemented several
\textsc{Closest} and \textsc{HiddenBy} predicates: If $p=\{1,2\}$,
exact computations are proposed and all predicates are in $O(1)$ with
only integer number computations
\cite{Hirata,Maurer2003,roerdnik}; If $p\in \R$, $p\geq 1$, we
have approximated computations on real numbers (\emph{double}) and we
consider the Generic \textsc{HiddenBy} predicate in $O(\log N)$
(Alg.~\ref{alg:genhidden}). Since predicates are based on floating
point computations, numerical issues may occur. If $p\in \Z$, $p\geq
3$, we use exact integer number based computations of distances
storing sum of power $p$ quantities (which can be computed in
$O(\log{p})$ thanks to exponentiation by squaring). The
\textsc{HiddenBy} predicate is also based on
Algorithm~\ref{alg:genhidden}.
Beside these predicates for $L_p$ metrics, \textsc{DGtal} also
contains a generic metric adapter: if the user specifies a distance
function (taking two points and returning a value) corresponding to a
norm with axis symmetric unit ball, generic \textsc{Closest} and \textsc{HiddenBy} predicates can be
automatically constructed. Please note that since all algorithms are
separable, the generic framework provided in \textsc{DGtal} allow us to
have a free multi-thread implementation \cite{dcoeurjo_ChapDTWADGMM}.
